\documentclass[english, LaM, oneside]{sapthesis}%remove "english" for a thesis written in Italian
%Bachelor's (laurea triennale) thesis : Lau 
%Master's (laurea specialistica) thesis: LaM 
%PhD's thesis: PhD 
%\usepackage[italian]{babel} %use this package for a thesis written in Italian
\usepackage[utf8]{inputenx}
\usepackage{indentfirst}
\usepackage{microtype}
%\usepackage{chemformula}
%\usepackage{setspace}
%\usepackage{yfonts,color}
%\usepackage{siunitx}
%\usepackage{comment}
%\usepackage{multirow}
%\usepackage{varioref}
%\usepackage[bottom]{footmisc}
%\usepackage{wrapfig}
%\usepackage{float}
%\usepackage{type1cm}
%\usepackage{lettrine}
\usepackage{setspace}
\singlespacing
%\linespread{0.9}
%\usepackage{chngcntr}
\usepackage[nottoc, notlof, notlot]{tocbibind}
%\onehalfspacing
%\counterwithout{footnote}{chapter}
\usepackage{hyperref}
\hypersetup{
			hyperfootnotes=true,			
			bookmarks=true,			
			colorlinks=true,
			linkcolor=red,
            linktoc=page,
			anchorcolor=black,
			citecolor=red,
			urlcolor=blue,
			pdftitle={Parametrized Counterfactual Explanations in Graph Neural Networks},
			pdfauthor={Giammarco D'Alessandro},
			pdfkeywords={thesis, sapienza, roma, university}
 }

\title{Parametrized Counterfactual Explanations in Graph Neural Networks: Interpreting Discrete Latent Structures for eXplainable AI}
\author{Giammarco D'Alessandro}
\IDnumber{1753102}
\course[]{Engineering in Computer Science - Ingegneria Informatica}
\courseorganizer{Facolt\`a di Ingegneria dell'Informazione, Informatica e Statistica}
\submitdate{2022/2023}
\copyyear{2023}
\advisor{Prof. Fabrizio Silvestri}
\coadvisor{Prof. Simone Scardapane}
\authoremail{dalessandro.1753102@studenti.uniroma1.it}
\examdate{... October 2023}
\examiner{Prof. ...} \examiner{Prof. ...} \examiner{Prof. ...}  \examiner{Prof. ...}  \examiner{Prof. ...} \examiner{Prof. ...}  \examiner{Prof. ...} 

%we refer to http://ctan.mirrorcatalogs.com/macros/latex/contrib/sapthesis/sapthesis-doc.pdf for an exhaustive description of the sapthesis documentclass.


\begin{document}

\frontmatter
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%% Abstract
\begin{abstract}
write your abstract here
\end{abstract}

\tableofcontents

\mainmatter
%%%%%%%%%%%%%%%%%%%%%%%%% CHAP.1 : Introduction 
\chapter{Introduction}
\label{chap:1} 
\lettrine[lines=2, findent=3pt, nindent=0pt]{I}{}n the frame of astronomical spectroscopy...

\bigskip
In \hyperref[chap:1]{Chapter~\ref*{chap:1}} we  briefly present...

\bigskip
In \hyperref[chap:2]{Chapter~\ref*{chap:2}} we summarize...

\section{Graph Neural Networks}
\label{sec:gnn}
\subsection{why are so popular}
\subsection{how they work}

\section{Explainability - XAI}
\label{sec:xai}
\subsection{Explainability in general}
\subsection{Explainability in GNNs}
\subsection{Counterfactual reasoning}



%%%%%%%%%%%%%%%%%%%%%%%%% CHAP.2 : My method
\chapter{Counterfactual Parametrized Explainer for Graph Neural Networks}
\label{chap:2} 
\section{The Problem of global explanations}
\subsection{Discrete Latent Structures}
\label{sec:caso}
Over the last decades, innovative observational techniques have been developed to allow spectrographs observing...


%%%%%%%%%%%%%%%%%%%%%%%%% CHAP.3 : Experiments
\chapter{Experimental results}
\label{chap:3}
\section{The Very Large Telescope}
Property of the European Southern Observatory...

\section{The Multi-Object Optical and Near-infrared Spectrograph}
\label{sec:moons}

The \textit{Multi-Object Optical and Near-infrared Spectrograph} is a future generation MOS instrument for the VLT. 

%%%%%%%%%%%%%%%%%%%%%%%%% CHAP.4 :
\chapter{Conclusions}
\label{chap:4} 
The grasping power of the mirror..



%%%%%%%%%%%%%%%%%%%%%%%%% bibliography
\backmatter
\phantomsection
\begin{thebibliography}{17}

\bibitem{ref:vph}
Blanche P.A., Gailly P., et al., “\textit{Volume phase holographic gratings: large size and high diffraction efficiency}“, Optical Engineering, Vol. 43, No.11, November 2004

\bibitem{ref:science}
Cirasuolo M., et al., \textit{"MOONS Science Report"}, MOONS Document Number: VLT-TRE-MON-14620-0001, Issue: $1.0$, $31^{\textup{st}}$ January $2013$

\bibitem{ref:eso}
European Southern Observatory, \url{http://www.eso.org}

\end{thebibliography}

\end{document}